---
title: "Example report for PM2.5"
output: html_notebook
---

```{r}
source('R/configuration.R')
source('R/load_functions.R')
source('R/render_functions.R')
source('R/population_functions.R')
```

__Hint__
The configuration assumes that ceipr data is present at 'webdab1.umweltbundesamt.at/download/gridding2018/' relative to this file. If you have downloaded it elsewhere, simply override this variable.


```{r}
neighbour_filter <- function(tbl) filter(tbl,iso2=="BE"|iso2=="NL"|iso2=="DE"|iso2=="GB"|iso2=="FR")
pollution_trend_all_sectors('CO',neighbour_filter)
```

## Sectors
These emission sectors are available:

```{r}
# This d
knitr::kable(sectors,title="Sectors")
```

Here's a helpful map of EU countries as well:

![](https://upload.wikimedia.org/wikipedia/commons/8/86/Europe_ISO_3166-1.svg)


# Example 1: Pollution trend per country

This loads the data for a specific pollutant and sector. As every zipfile contains data from _all_ EU countries, this would be the (only) logical variable to group by. Since we don't specify a year, all years will be loaded.

```{r}
# Read the data for CO in sector "H" (aviation), all years will be loaded:
NOx_pollutant <- 'NOx'
H_sector <- 'F'
NOx_data <- read_pollution_data(pollutant=NOx_pollutant,sector=H_sector)
# Define a filter to select countries of interes:
belgian_neighbours_filter <- function(tbl) filter(tbl,iso2=="BE"|iso2=="NL"|iso2=="DE"|iso2=="GB"|iso2=="FR")
view_trends(belgian_neighbours_filter(NOx_data),subject=paste(NOx_pollutant,sectors[H_sector]))
```

# Example 2: Pollution trends per sector for one country

This loads the data for a specific pollutant for all sectors. As every zipfile contains data from _all_ EU countries, we'll need to narrow down to one specific country.

```{r}
# Read the data for NOx for all sectors and all years:
pollutant_name <- 'NOx'
pollution_data <- read_pollution_data(pollutant = pollutant_name)
# Define a filter to select countries of interes:
belgian_filter <- function(tbl) filter(tbl,iso2=="BE")
view_trends(belgian_filter(pollution_data),group="sector",group_label = "Belgian sector",subject = pollutant_name)
```

# Example 3: Comparison of totals per country in a year

This loads the data for a specific pollutant and sector. As every zipfile contains data from _all_ EU countries, this would be the (only) logical variable to group by. Since we're only comparing the totals for one year, we can view all countries.

```{r}
# Read the data for CO in sector "H" (aviation), all years will be loaded:
name_pollutant <- 'NOx'
name_sector <- 'F'
selected_year <- 2016
pollution_data <- read_pollution_data(pollutant=name_pollutant,sector=name_sector,year=selected_year)
# Since the data will contain a lot of countries, we subselect Eurozone countries.
view_comparisons(filter(pollution_data,iso2 %in% euro_zone))
```

# Example 4: Rank countries by pollution per person

```{r}
# Read the data for CO in sector "F" (road transport), all years will be loaded:
name_pollutant <- 'NOx'
name_sector <- 'F'
selected_year <- 2016
pollution_data <- read_pollution_data(pollutant=name_pollutant,sector=name_sector,year=selected_year)
pollution_data <- filter(pollution_data,iso2 %in% euro_zone)
pollution_totals <- dplyr::group_by(pollution_data,iso2,year) %>%
    dplyr::summarise(total = sum(emission))
pollution_totals <- add_population(pollution_totals)
pollution_totals$emission_per_person <- with(pollution_totals,1000*total/population)
# Since the data will contain a lot of countries, we subselect Eurozone countries.
# view_comparisons(filter(pollution_data,iso2 %in% euro_zone))
ggplot(pollution_totals,aes(x=reorder(iso2,-emission_per_person),y=emission_per_person,fill=iso2)) +
   geom_bar(stat="identity",position="dodge")
```
